---
title: "Open OGC Geoservices in Switzerland"
subtitle: "Exploratory Analysis of the available OGC services"
format:
  html:
    embed-resources: true
editor: visual
execute:
  echo: false
  warning: false
  message: false
---

## OGC Webservices in Switzerland

This document explores the OGC webservices provided by the various public open data provider. These OGC webservices are publicly available and provide broad and well maintained metadata of the content they portray.

This analysis is based on the harvested OGC web services (WMS, WMTS, WFS) of Switzerland by the following repository: <https://github.com/davidoesch/geoservice_harvester_poc> by [David Oesch](https://github.com/davidoesch). The *geoservice harvester* crawls the known OGC web services managed by the state, cantons and communes.

```{r directorySetup,comment=FALSE, message = FALSE, echo=FALSE,warning=FALSE}
# rm(list=ls())         # Clean the environment (comment out, if batch script)
options(scipen=999)     # display digits proberly!! not the scientific version
options(digits.secs=6)  # use milliseconds in Date/Time data types
options(warning=FALSE)  # don't show warnings
library(knitr)          # set global knitr options of the document
# memory.limit(size=1800) # set memory limit 

# Libraries
library(rgdal)# install.packages("rgdal") # install libraries
library(dplyr)
library(tidyr)
library(sf)
library(ggplot2)
library(ggrepel) # https://github.com/slowkow/ggrepel/issues/89
library(RColorBrewer)
library(igraph)
library(kableExtra) # http://haozhu233.github.io/kableExtra/best_practice_for_newline_in_latex_table.pdf
options(kableExtra.auto_format = TRUE)
library(tm)


# Folders
dataFolder   <- file.path("data")   # Set path to the data and figure folder
resultFolder <- file.path("result") # Set path to the data and figure folder
# Settings
themeReport <- theme_minimal() # customise plot theme
colorPal <- c("#ddf1da","#abdda4","#e6f598","#fee08b","#fdae61","#f46d43","#d53e4f") # set a color palette for the map
colorPrimary <- "#1f78b4"
caption <- "Data Source: OGC Webservices Switzerland"
```

```{r readdata}
# Set file paths
geoCHPath = "https://raw.githubusercontent.com/davidoesch/geoservice_harvester_poc/main/data/geoservices_CH.csv"
geoSimpleCHPath = "https://raw.githubusercontent.com/davidoesch/geoservice_harvester_poc/main/data/geodata_simple_CH.csv"
geoservicesCHPath = "https://raw.githubusercontent.com/davidoesch/geoservice_harvester_poc/main/data/geoservices_CH.csv" # TODO chec if duplicate of geoCHPath

# read files
geoCH <- read.csv(geoCHPath, stringsAsFactors = FALSE, encoding="UTF-8")
# geoSimpleCH <- read.csv(geoSimpleCHPath, stringsAsFactors = FALSE, encoding="UTF-8")
# geoservicesCH <- read.csv(geoservicesCHPath, stringsAsFactors = FALSE, encoding="UTF-8")

# Read Spatial data
cantonShapeFile <- file.path(dataFolder,"BFS_GGG/k4k20.shp")
lakeShapeFile <- file.path(dataFolder,"BFS_GGG/k4s20.shp")
cantonSF <- st_read(dsn = cantonShapeFile,stringsAsFactors = FALSE,quiet=TRUE)
lakesSF<- st_read(dsn = lakeShapeFile,stringsAsFactors = FALSE,quiet=TRUE)

```

### Dataset

| Dataset               | Description                                                     |
|---------------------|---------------------------------------------------|
| geoservices_CH.csv    | OGC Services of Switzerland full dataset                        |
| geodata_simple_CH.csv | OGC Services of Switzerland reduced dataset (Owner, Title, Url) |

```{r}
id <- sample.int(nrow(geoCH),1) # random service
d <- data.frame(key=names(geoCH),example=as.character(geoCH[id,]))
knitr::kable(d, digits = 2,row.names=FALSE, col.names = c("Key","OGC Service Example"),caption = "Field names and OGC Service Value Example", booktabs = TRUE,caption.short = "Random OGC Service") %>% column_spec(2,width="8 cm") %>%  kable_classic(font_size = 12)
```

**Spatial Datasets**

The Federal Statistical office BFS provides generalized boundaries of local and regional authorities. The data sets are particularly suitable for visualising statistical data on communes, districts and cantons as well as for creating small scale maps relating to different topics.

BFS Generalisierte Gemeindegrenzen der Schweiz \[[BFS Generalisierte Gemeindegrenzen der Schweiz 2020](https://www.bfs.admin.ch/bfs/de/home/dienstleistungen/geostat/geodaten-bundesstatistik/administrative-grenzen/generalisierte-gemeindegrenzen.assetdetail.14776419.html),[Dataset](https://www.bfs.admin.ch/bfsstatic/dam/assets/14776419/master)\]

### Distribution of Content by Service Owners

How many layers do the various service provider offer? How are the services structured?

```{r}
d <- geoCH %>% group_by(OWNER) %>% summarise(layers = length(unique(NAME)), links = length(unique(SERVICELINK)))
```

```{r}
#| fig-cap: Number of Layers provided by each service owner 
#| fig-height: 6
ggplot(data = d, aes(x=reorder(OWNER,layers), y= layers)) + 
  geom_bar(stat = "identity",fill=colorPrimary) +
  geom_text(aes(label = layers), nudge_y = 2,size=3, color = "black",hjust = -0.2)+
   scale_y_continuous(expand = expansion(mult = c(0, .08)))+
  coord_flip() +
   theme(axis.line.x = element_line(size = .2),panel.background = element_blank(),axis.ticks.y=element_blank())+
  labs(title = "Distribution of Content by Service Owners", subtitle= "Number of layers by Service Owner",x = NULL,y = "Number of Layers",fill="",caption=caption)
```

```{r}
#| fig-cap: Number of unique service links provided by each service owner 
#| fig-height: 6
ggplot(data = d, aes(x=reorder(OWNER,links), y= links)) + 
  geom_bar(stat = "identity",fill=colorPrimary) +
  geom_text(aes(label = links), nudge_y = 2,size=3, color = "black",hjust = -0.2)+
   scale_y_continuous(expand = expansion(mult = c(0, .08)))+
  coord_flip() +
   theme(axis.line.x = element_line(size = .2),panel.background = element_blank(),axis.ticks.y=element_blank())+
  labs(title = "Distribution of Service Links", subtitle= "Number of Service Links by Service Owner",x = NULL,y = "Number of Service Links",fill="",caption=caption)

```

```{r}
#| fig-cap: Number of unique service links provided by each service owner 
#| fig-height: 6

ggplot(data = d, aes(x=layers, y= links)) + 
  geom_point(size=1,color=colorPrimary)+
  # geom_text(aes(label = OWNER), nudge_y = 2,size=3, color = "black",hjust = -0.2)+
  geom_text_repel(mapping = aes(x = layers,y = links,label = OWNER),size = 2.4,min.segment.length = 0,point.padding = 0.5,segment.color = "grey30", max.overlaps=13)+ scale_y_continuous(trans = "log10")+
  themeReport +labs(x = "Number of Layers",y = "Number of Service Links",title = "Number of Service Url vs Layers", caption=caption)

```

## Spatial Coverage of Geoservices

```{r}
# create bounding boxes from services

# convert bounding box string to a polgon 
# eg. "2740443.137 1232276.032 2765413.73 1258037.171 EPSG:2056"
# https://github.com/r-spatial/sf/issues/1034
str2BBox <- function(bboxString){
  i = 0
  res <- lapply(bboxString,function(str){
    i <- i +1
    # extract coordinates and EPSG Code
    x <- unlist(strsplit(str ," "))
    x <- x[x != ""] # remove empty strings
    # return an empty polygon if coordinates are missing
    if(length(x)==0){return(st_polygon())} 
    # epsg <- as.numeric(unlist(strsplit(x[5] ,":"))[2])
    x <- as.numeric(x[1:4])
    if(anyNA(x)){cat(i,": ",x,"has NA")}
    # create a bounding box polygon
    x <- st_polygon(list(cbind(
          c(x[1],x[3],x[3],x[1],x[1]), # x coords 
          c(x[2],x[2],x[4],x[4],x[2])  # y coords
          )))
    return(x)
  })
  return(res)
}

# extract coordinates and EPSG Code
# eg. "2740443.137 1232276.032 2765413.73 1258037.171 EPSG:2056"
str2EPSG <- function(bboxString){
  res <- lapply(bboxString,function(str){
    epsg <- as.numeric(unlist(strsplit(str,":"))[2])
    return(epsg)
  })
  return(unlist(res))
}
```

```{r}
# generate sf bounding boxes
geoBBox <- str2BBox(geoCH[,"BBOX"])
geoEPSG <- str2EPSG(geoCH$BBOX)
geoBBoxStr <- lapply(geoBBox,st_as_text)

geoCHSF <- st_sf(geoCH,geoBBox)
geoCHSF$epsg <- str2EPSG(geoCH$BBOX)
geoCHSF$isEmpty <- st_is_empty(geoCHSF) # which polygons have no geometries

geoCH$epsg <- str2EPSG(geoCH$BBOX)
geoCH$isEmpty <- st_is_empty(geoCHSF) # which polygons have no geometries
# geoCH$bboxStr <- lapply(geoBBox,st_as_text)
```

```{r}
# EPSG Count
d <- geoCH %>% group_by(epsg) %>% summarise(epsg = first(unique(epsg)), n = n())
knitr::kable(d, digits = 2,row.names=FALSE,caption = "Frequency of CRS provided for Geoservies in CH (epsg)", booktabs = TRUE,caption.short = "EPSG")  %>% column_spec(2,width="8 cm") %>%  kable_classic(font_size = 12)
```

```{r}
#| fig-cap: Frequency of CRS provided for Geoservies in CH (epsg)
#| fig-height: 2

ggplot(data = d, aes(x=reorder(epsg,n), y= n)) + 
  geom_bar(stat = "identity",fill=colorPrimary) +
  geom_text(aes(label = n), nudge_y = 2,size=3, color = "black",hjust = -0.2)+
   scale_y_continuous(expand = expansion(mult = c(0, .08)))+
  coord_flip() +
   theme(axis.line.x = element_line(size = .2),panel.background = element_blank(),axis.ticks.y=element_blank())+
  labs(title = "Distribution of CRS by Dataset", subtitle= "Frequency of EPSG Occurence",x = NULL,y = "Count",fill="",caption=caption)

```

```{r}
dsf <- geoCHSF %>% filter(epsg==2056)
st_crs(dsf) <- 2056
ggplot() + themeReport  +
  geom_sf(data = cantonSF,fill=NA, colour="#cccccc",size=0.25) +
  geom_sf(data = lakesSF, fill="#a6cee3", color="#a6cee3",alpha=0.6,size=0.25) +
  geom_sf(data = dsf,fill=NA, colour="#ff660020",size=0.25) +
  labs(title = "GeoServices Switzerland",subtitle = "Spatial Coverage of the Bounding Boxes with EPSG 2056",fill = "",x = NULL,y = NULL,colour="", caption=paste(caption,", BFS",sep=""))+coord_sf(datum = NA)
```

```{r}

dsf <- geoCHSF %>% filter(epsg==2056)
st_crs(dsf) <- 2056
# bbox of switzerland
# st_bbox(cantonSF)
# 2485202 1075281 2834036 1295864 
# 2400000 1000000 2850000 1300000 


box = c(xmin = 2400000, ymin = 1000000, xmax = 2850000, ymax = 1300000)
dsf <- st_crop(dsf,box )


ggplot() + themeReport  +
  geom_sf(data = cantonSF,fill=NA, colour="#cccccc",size=0.25) +
  geom_sf(data = lakesSF, fill="#a6cee3", color="#a6cee3",alpha=0.6,size=0.25) +
  geom_sf(data = dsf,fill=NA, colour="#ff660020",size=0.25,alpha=0.01) +
  labs(title = "GeoServices Switzerland",subtitle = "Spatial Coverage of the Bounding Boxes with EPSG 2056",fill = "",x = NULL,y = NULL,colour="", caption=paste(caption,", BFS",sep=""))+coord_sf(datum = NA)

```

```{r}
# generate word cloud from service title, abstract, keywords
library(tm)
library(wordcloud)
library(RColorBrewer)
# function to create a clean text corpus
createCorpus <- function(text){
  docs <- Corpus(VectorSource(text))
  docs <- docs %>%
    tm_map(removeNumbers) %>%
    tm_map(removePunctuation) %>%
    tm_map(stripWhitespace)
  docs <- tm_map(docs, content_transformer(tolower))
  
  docs <- tm_map(docs, removeWords, stopwords("english"))
  docs <- tm_map(docs, removeWords, stopwords("german"))
  docs <- tm_map(docs, removeWords, stopwords("french"))
  docs <- tm_map(docs, removeWords, stopwords("italian"))
  return(docs)
}
createTermDocumentMatrix <- function(docs){
  dtm <- TermDocumentMatrix(docs) 
  matrix <- as.matrix(dtm) 
  words <- sort(rowSums(matrix),decreasing=TRUE) 
  df <- data.frame(word = names(words),freq=words)
  return(df)
}

# create a corpus with data cleaning
ctitle <- createCorpus(geoCH$TITLE)
cabstract <- createCorpus(geoCH$ABSTRACT)
ckeywords <- createCorpus(geoCH$KEYWORDS)

# create a TermDocumentMatrix
tdmtitle <- createTermDocumentMatrix(ctitle)
tdmabstract <- createTermDocumentMatrix(cabstract)
tdmkeywords <- createTermDocumentMatrix(ckeywords)
```

```{r,warning=FALSE}
# create wordcloud
df <- tdmtitle
wordcloud(words = df$word, freq = df$freq, min.freq = 10, max.words=200, random.order=FALSE, rot.per=0.35, colors=brewer.pal(9, "Blues")[3:9])
```

```{r,warning=FALSE}
# create wordcloud
df <- tdmabstract
wordcloud(words = df$word, freq = df$freq, min.freq = 10, max.words=200, random.order=FALSE, rot.per=0.35, colors=brewer.pal(9, "Blues")[3:9])
```

```{r,warning=FALSE}
# create wordcloud
df <- tdmkeywords
wordcloud(words = df$word, freq = df$freq, min.freq = 10, max.words=200, random.order=FALSE, rot.per=0.35, colors=brewer.pal(9, "Blues")[3:9])
```

```{r}
#| fig-cap: Most frequent Terms for Geoservice title 
#| fig-height: 8
d <- tdmkeywords[1:50,]
ggplot(data = d, aes(x=reorder(word,freq), y= freq)) + 
  geom_bar(stat = "identity",fill=colorPrimary) +
  geom_text(aes(label = freq), nudge_y = 2,size=3, color = "black",hjust = -0.2)+
   scale_y_continuous(expand = expansion(mult = c(0, .08)))+
  coord_flip() +
   theme(axis.line.x = element_line(size = .2),panel.background = element_blank(),axis.ticks.y=element_blank())+
  labs(title = "Geoservice Titles in Switzerland", subtitle= "Term Frequency",x = NULL,y = "Count",fill="",caption=caption)
```

```{r}
#| fig-cap: Most frequent Terms for Geoservice abstract 
#| fig-height: 8
d <- tdmabstract[1:50,]
ggplot(data = d, aes(x=reorder(word,freq), y= freq)) + 
  geom_bar(stat = "identity",fill=colorPrimary) +
  geom_text(aes(label = freq), nudge_y = 2,size=3, color = "black",hjust = -0.2)+
   scale_y_continuous(expand = expansion(mult = c(0, .08)))+
  coord_flip() +
   theme(axis.line.x = element_line(size = .2),panel.background = element_blank(),axis.ticks.y=element_blank())+
  labs(title = "Geoservice abstracts in Switzerland", subtitle= "Term Frequency",x = NULL,y = "Count",fill="",caption=caption)
```

```{r}
#| fig-cap: Most frequent Terms for Geoservice keywords 
#| fig-height: 8
d <- tdmkeywords[1:50,]
ggplot(data = d, aes(x=reorder(word,freq), y= freq)) + 
  geom_bar(stat = "identity",fill=colorPrimary) +
  geom_text(aes(label = freq), nudge_y = 2,size=3, color = "black",hjust = -0.2)+
   scale_y_continuous(expand = expansion(mult = c(0, .08)))+
  coord_flip() +
   theme(axis.line.x = element_line(size = .2),panel.background = element_blank(),axis.ticks.y=element_blank())+
  labs(title = "Geoservice keywords in Switzerland", subtitle= "Term Frequency",x = NULL,y = "Count",fill="",caption=caption)
```
